Negative and Fractional Types

### 2012-06-27 rational_ext.tex

%%%%%%%%%%%%%%%%%
\subsection{Vector Spaces}

In order to interpret fractional types, we first lift the values and
combinators of {{Pi}} to the world of vector spaces. Intuitively:
\begin{itemize}
\item each type {{b}} is now interpreted as a vector space of dimension
  $\modulusp{b}$ over the field $\Fpx{2}$; the elements of such a vector
  space are column vectors
\item each value {{v_i : b}} is interpreted as column vector:
\[\begin{array}{r@{~}l}
\begin{array}{r} 
v_0 \\ v_1 \\ \vdots \\ v_i \\ \vdots \\ v_n 
\end{array} & 
\begin{bmatrix}
0 \\ 0 \\ \vdots \\ 1 \\ \vdots \\ 0
\end{bmatrix}
\end{array}\]
\item each combinator {{c : b1 <-> b2}} is interpreted as a matrix whose
  columns are indexed by the values of type {{b1}} and whose rows are indexed
  by the values of type {{b2}}. The matrix is \emph{orthogonal} with each row
  and column containing exactly one occurrence of the value 1.
\end{itemize}

Once we are in the world of vector spaces, the fractional type {{1/b}} is
interpreted as the \emph{dual} to the vector space denoted by {{b}}. The
elements of this dual vector space can be represented as \emph{row vectors}.

Multiplying a row vector by a column gives a scalar: this is eta and epsilon.

>>>>

%%%%%%%%%%%%%%%%%
\subsection{Pointed Sets}

In order to interpret fractional types, we first lift the values and
combinators of {{Pi}} to the world of \emph{pointed sets}. Intuitively:
\begin{itemize}
\item instead of interpreting a type {{b}} as a plain set of cardinality
  $\modulusp{b}$, we now interpret each type {{b}} as a pointed set $P_b$
  with a distinguished \emph{basepoint} $\bot_b$;
\item a combinator {{c : b1 <-> b2}} is interpreted as a function as before
  with the restriction that it must map the basepoint $\bot_{b_1}$ to the
  basepoint $\bot_{b_2}$.
\end{itemize}

Intuitively a pointed set may be seen as a vector space over the field with
one element. This view allows us to add the type {{1/b}} and interpret it as
the \emph{dual} of the vector space denoting {{b}}. Concretely the elements
of the pointed set denoted by {{1/b}} are \emph{linear functionals} $f : P_b
\rightarrow P_1$. Note that $P_1$ is the pointed set containing {{()}} and
$\bot_1$. The isomorphism between the types {{1}} and {{b * 1/b}} is now a
consequence of that fact that the pairing of a vector and a functional in the
dual space produces a scalar.
